Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen...' 

### 2017-02-22 salmon/paper/salmon_paper_appendix_v3.Rmd
Revert "Revert "test""
This reverts commit 5085fdd090082f92c5e3c1796c48806f08a6e80d.

---
title: ""
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

@@

```{r, echo = FALSE, eval=FALSE}
dat = dat.salmon
# create other category, lumping coho/chinook
dat$salm = dat$coho+dat$chnk
dat$salm.prev = dat$coho.prev+dat$chnk.prev

dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev

dat$salm = dat$salm/dat$totsalm
dat$sock = dat$sock/dat$totsalm
dat$pink = dat$pink/dat$totsalm
dat$chum = dat$chum/dat$totsalm
dat$chnk = dat$chnk/dat$totsalm
dat$coho = dat$coho/dat$totsalm
dat$salm.prev = dat$salm.prev/dat$totsalm.prev
dat$sock.prev = dat$sock.prev/dat$totsalm.prev
dat$pink.prev = dat$pink.prev/dat$totsalm.prev
dat$chum.prev = dat$chum.prev/dat$totsalm.prev
dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
dat$coho.prev = dat$coho.prev/dat$totsalm.prev

dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
dat$logdiff = log(dat$revenue/dat$revenue.prev)

dat_old = dat
dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
  abs(dat$pink-dat$pink.prev) + 
  abs(dat$chum-dat$chum.prev) + 
  abs(dat$sock-dat$sock.prev)
dat = dat[which(dat$sumP2 < 0.3),]
```

```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere"}

df = dat[which(dat$strategy_permit=="S01E"),]
pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")

g = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T))

gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
```

@@

```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2, eval=FALSE}
### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)

daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
  summarize(p = sum(g_earn)) %>%
  group_by(p_holder, year) %>% mutate(p = p/sum(p))
daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
daty[is.na(daty)]=0
ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
  geom_point(alpha=0.3,size=1) +
  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
```

```{r, echo=FALSE, eval=FALSE}

mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)

cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
# quadratic model, 
# 'squared term' indicates concave up/down
# derivative is magnitude, 2a * x + b

df = ranef(mu.mod)$cond$year_permit
colnames(df) = c("c","b","a")
df$year = as.numeric(substr(rownames(df), 1, 4))
df$strategy = substr(rownames(df), 6, 9)
df$deriv = 2*df$a + df$b
# also calculate change of going from 1 to 2 spp
df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
# add fixed effects into predictions

# We can demonstrate the quadratic curves by year for a single strategy / set of years
df.s01e = df[df$strategy=="S01E",]
demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
demo = left_join(demo, df.s01e)
demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
```

```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide', eval=FALSE}
pdf("../../salmon/Salmon_residuals.pdf")
# Residuals for mean model
dat.salmon$mu.resid = residuals(mu.mod)
dat.salmon$mu.fitted = fitted(mu.mod)

ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")

# Residuals for CV model
dat.salmon$cv.resid = residuals(cv.mod)
dat.salmon$cv.fitted = fitted(cv.mod)

ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
dev.off()
```
```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. ", eval=FALSE}
###We can make quadratic plots of these effects by year and strategy. These plots represent the effect of increasing diversity (as relative change) on mean revenue. For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more. In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.

ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3, eval=FALSE}
###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Across permits, perhaps slight declines in some of these in recent years?

ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3, eval=FALSE}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit.", eval=FALSE}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")

```{r, echo=FALSE, eval = FALSE}
# These are the run  timing plots per Milo's questions

# for 2008, create cumulative density plots of salmon landings by calday
pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]

# aggregate by calendar day
pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)

group_by(pinkchum, jday, year) %>% 
  summarize(pink = sum(g_earn[spec=="PINK"]), 
    chum = sum(g_earn[spec=="CHUM"]),
    sock = sum(g_earn[spec=="SOCK"])) %>% 
  group_by(year) %>% 
  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")

# make plot of all earnings or weight by jday
salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
salm$month = as.numeric(substr(salm$day, 1, 2))
salm$day = as.numeric(substr(salm$day, 4, 5))
salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)

g = group_by(salm, jday, year) %>% 
  summarize(earn = sum(g_earn), 
    lbs = sum(g_pounds), 
    pink = sum(g_earn[spec=="PINK"])/earn,
    sock = sum(g_earn[spec=="SOCK"])/earn,
    chum = sum(g_earn[spec=="CHUM"])/earn,
    coho = sum(g_earn[spec=="COHO"])/earn,
    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", fig.pos="placeHere", eval=FALSE}

df = dat[which(dat$strategy_permit=="S01A"),]
pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

\pagebreak

@@

```{r, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
appendixPlot1
```

@@

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy,scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue") + xlab("Year")
```

```{r, echo=FALSE, fig.height = 3.1, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons / person") + xlab("Year") + ggtitle("Trends in landed weight")
```

@@

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
```

## Species diversity

```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3, eval=FALSE}
## Species composition
# plot proportion of species landed by salmon permit
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
  ylab("Species % (total) ") + xlab("")
```

```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
group_by(dat.salmon, year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)") + xlab("")
```

```{r, fig.cap="Proportion of total returns to Prince William Sound (hatchery + wild, catch + escapement). Years before 1980 and after 2011 aren't included because not all species have data.", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
returns = readxl::read_excel("salmon_returns_PWS.xlsx")
returns = returns[,c("Year","Sockeye_total","Pink_total","Chum_total","Coho","Chinook")]

returns = as.data.frame(returns[which(returns$Year%in%c(1980:2011)),])
for(i in 1:ncol(returns)) returns[,i] = as.numeric(returns[,i])
for(i in 1:nrow(returns)) {
    # normalize
    returns[i,-1] = returns[i,-1] / sum(returns[i,-1])
}

melted = reshape2::melt(returns,id.vars="Year")
names(melted)[which(names(melted)=="variable")]="species"
ggplot(melted,aes(Year,value, group = species, color=species)) + geom_line() + ylab("Proportion of total returns (by #)")
```

>>>>

---
title: "Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen...' "
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

@@

@@

@@

```{r figa1, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
print(appendixPlot1)
```

@@

```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
```

```{r, echo=FALSE}
#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
#cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))

# plot proportion of species landed by salmon permit
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

#cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
#cfec.raw = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]

g1 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01A") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g2 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g3 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01K" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g4 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01M" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g5 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g6 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec.raw$year > 0,], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

dat.diff = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

```{r figa3, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit for seine fisheries in Southeast Alaska (S01A), Prince William Sound (S01E) and Kodiak (S01K). Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)", fig.height=3.2}
gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
```

```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit, for seine fisheries on the Alaska Peninsula (S01M), and gillnet fisheries in Southeast Alaska (S03A) and Prince William Sound (S03E).", fig.height=3.2} 
gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
```

@@

## Model Results

```{r figa5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}

diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))
# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)
yearpermitold=yearpermit

# grab original data to look at permit names
diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = rstan::extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev
x3 = log(specdiv/specdiv.prev) * specdiv.prev^2

df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0

for(i in 1:nrow(df)) {
  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
  tmp = 
    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
  df$pred[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}

ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_line() +
  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")

```{r figa6, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit. These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
# grab original data to look at permit names
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")

years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))

# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

yearpermitold=yearpermit
load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold
est = rstan::extract(mod, permute="TRUE")

yearpermit$b0 = apply(est$b0_str_yr,2,median)
yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
